### 2023-02-04 src/tex/ms.tex
finalized most of the figures

% Define document class
\documentclass[preprint]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

\newcommand{\ergsec}{erg s$^{-1}$ cm$^{-2}$ $\mu$m$^{-1}$\ }

% Title
\title{Keck Near-Infrared Observations of Mab}

% Author list
\author{@emolter}

\author{Imke}

\author{Chris}

\begin{abstract}
We report the first near-infrared detection of Uranus's smallest known moon Mab, the presumed source of the blue and diffuse $\mu$ ring, using the NIRC2 instrument at Keck Observatory.
something about shift-and-stack
If the moon has an effective radius of 12 km as previously proposed based on HST data, its geometric albedo would be xx at 1.6 $\mu$m and phase angle 0.02$^\circ$, while if it has a radius of 6 km, its geometric albedo would be xx. Taken along with the HST albedo measurements, this implies Mab is (much fainter than / much brighter than / about the same as) expected in the near-infrared. We propose that this is due to ... (remember to talk about low phase angle and opposition surge somewhere). We also report (the first?) infrared photometry of Ophelia, (Cordelia?), and Perdita.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{s:intro}

something about the circumplanetary environment of Uranus

% does the below text fit in anywhere?
The small regular moons comprising the Portia group in the Uranian system represent the most densely-packed set of satellites known, with semimajor axis separations of 3\% on average \citep{showalter20}; as such, the system is predicted to be chaotic, with collisions occurring on 10$^5$-10$^7$-year timescales \citep{duncan97, french12}. Indeed, the faint dusty $\nu$ ring, situated between the orbits of Portia and Rosalind, has been proposed to be the remnant of such a collision. Due in part to this small orbital spacing, the Uranian satellites affect each other's geology via dust transport in the circumplanetary environment. For example, dust deposition from Uranus's outer irregular satellites has been invoked to explain spectral reddening of the leading hemispheres of Ariel, Umbriel, Titania, and Oberon \citep{cartwright18}. Material transport is important for understanding the surfaces of Miranda, Puck, and the smaller inner moons, too, but these processes are less well-studied. 
Puck's leading hemisphere has a significantly bluer spectral slope than its trailing hemisphere, meaning that it may be sweeping up material from the $\mu$ ring \citep{french17}, but this has not been confirmed. %this is a DPS abstract - is there a better reference?
Unusually low depth/diameter ratios of craters on the potential ocean world Miranda suggest the deposition of a $\sim$1 km thick regolith \citep{cartwright21, beddingfield22}. Those authors propose three possible sources of the regolith material: deposition from the $\mu$ ring, plume deposits from a subsurface ocean, and/or giant impact ejecta. 
One sentence here linking these paragraphs

Uranus's tiny moon Mab is embedded in the faint, diffuse $\mu$ ring, the only ring in the solar system with a blue spectrum other than Saturn's E ring, which is sourced from Enceladus's cryovolcanism \citep{depater06}. The E~ring is blue due to its abundance of tiny ($\lesssim$1 $\mu$m) particles and paucity of larger grains \citep{showalter91, depater04, hillier07}. While active cryovolcanism on such a small body as Mab seems highly unlikely, alternative theories for the formation and maintenance of the $\mu$ ring remain unconstrained by data. Mab has been proposed to be outbursting like the active Centaur Echeclus, with the blue color of the $\mu$ ring attributable not to small particle sizes but to a high abundance of carbon-rich organic matter \citep{seccull19}. Alternatively, interplanetary dust and/or charged plasma impacting Mab's surface may liberate enough particles to form the $\mu$ ring. To explain the blue color, the latter scenario relies on theoretical and numerical simulation expectations that $\sim$1 $\mu$m particles are more likely than larger particles to disperse into a ring (instead of re-accreting to the planet) under the combined influences of gravity, radiation pressure, and electromagnetic forces \citep{horanyi92, sfair09, sfair12}. In any case, the physical properties of this unusual moon are of interest for studies of interactions between airless surfaces and the interplanetary environment, the origin and dynamical evolution of circumplanetary dust, and the formation and evolution of small solar system bodies.

The radius and albedo of Mab remain under debate. Based on HST observations, \citet{showalter06} proposed two end-member cases: Mab has a $\sim$6 km radius with a Miranda-like icy surface and an albedo of $\sim$0.46; or Mab has a $\sim$12 km radius and an albedo of $\sim$0.1, indicating the presence of dark organic compounds on its surface like its other nearest neighbor, Puck. However, a sensitive near-infrared search with Keck did not detect the moon, leading to the proposal that Mab's radius is $\lesssim$6 km and as dark at wavelengths from 1.6-2.4 $\mu$m as Uranus's inner main rings \citep[albedo $\sim$0.05;][]{paradis19, paradis23}. Strong near-IR absorption from water ice could partially account for this darkening; however, the decrease in brightness is not expected to exceed 50\% even for a very icy body \citep[c.f. Haumea, one of the purest amorphous ice surfaces in the solar system;][]{gourgeot16}.

re-word the way Paradis is talked about above

Here we present the first observations of Mab at near-infrared wavelengths. Our dataset and data proceessing steps are described in Section \ref{s:observations}. We present photometry of Mab, as well as Cordelia, Ophelia, and Perdita,

\section{Observations and Data Processing}
\label{s:observations}

This paper makes use of H- and K-band data taken on 28 October 2019 by the narrow camera of the NIRC2 instrument coupled to the Adaptive Optics (AO) system at Keck Observatory, using Uranus itself for wavefront sensing \citep{wizinowich00, vandam04}. The same data were analyzed by \citet{paradis23}, so we refer the reader to that paper for a detailed description of the observation and data reduction procedures. The standard star was FS2 from the UKIRT Faint Standard Star list,\footnote{\url{https://www2.keck.hawaii.edu/inst/nirc/UKIRTstds.html}} and standard photometric calibration procedures resulted in calibration constants $C_H = 8.87\times10^{-17}$ and $C_K = 9.11\times10^{-17}$ to convert from 1 count per second to 1 \ergsec. The uncertainty in the photometric calibration was estimated from the spread in measured fluxes from the standard star between frames. The value we found was 20\% in both filters, which is typical for NIRC2 observations \citep{depater14, molter19}.

Our full dataset consisted of 23 exposures of 120 seconds each in H-band (1.6 $\mu$m) and 24 exposures of 120 seconds each in K-band (2.2 $\mu$m); these were all the frames taken on 28 October 2019 in which Mab's expected position was inside the field of view of the narrow camera; see the left panel of Figure \ref{fig:detection}.
The proper motion of Mab is roughly 30 mas (3 pixels) per minute, which means that each frame had to be shifted relative to the previous one in accordance with Mab's expected vector of motion to improve the signal-to-noise ratio (SNR). This so-called ``shift-and-stack" technique has been used by other authors to improve the SNR of moving point sources, including to measure photometry of other small moons of Uranus with NIRC2 \citep[][]{paradis19, paradis23}. Our shift-and-stack procedure differs in several ways from the one implemented by those authors, though, and we describe it fully in Appendix \ref{s:shiftandstack}. The application of shift-and-stack resulted in one stacked image per filter at the expected position of Mab, representing effective integration times of 2760 seconds in H-band and 2880 seconds in K-band.

\begin{figure}
\includegraphics[width=0.5\textwidth]{figures/motion_on_detector.png}
\includegraphics[width=0.5\textwidth]{figures/detection_images_Mab.png}
\caption{\textbf{Left:} A single NIRC2 H-band exposure overlaid with the expected position of Mab in each of our 23 H-band exposures. The bright and faint point sources visible in the image are Miranda and Puck, respectively. \textbf{Right:} Detection of Mab in H-band (top panel) and tentative detection in K-band (bottom panel) using the shift-and-stack technique.\label{fig:detection}}
\end{figure}

\section{Results}
\label{s:results}

\subsection{Photometry of Mab}

An image of our H-band detection of Mab is shown in the upper right panel of Figure \ref{fig:detection}. The moon appears slightly vertically extended because its proper motion was roughly 6 pixels per 120-second frame, moving approximately toward sky North. The detection is in a low-SNR regime, but we are confident that it is real for three reasons. First, a tentative detection of Mab can be made separately from the first 11 frames and the last 11 frames, meaning that an artefact in a single frame or 8-frame group is not producing a false positive. Second, the detection is located in the appropriate position on the sky relative to Miranda and Puck, as discussed below. Third, we performed an experiment to test the false positive rate of our shift-and-stack procedure under the same scattered light and noise conditions as the data. In short, we added random offsets to the expected position of Mab in each frame, then carried out shift-and-stack in the same way as we did for the true expected position. We repeated this procedure with different random offsets 100 times, and found that no noise spikes as bright as our Mab detection appeared anywhere in any of these 100 different stacks; see Appendix \ref{s:falsepositives} for more details.

Because Mab was detected at low SNR and within a substantial scattered light background, the flux in the wings of the PSF could not be accurately measured. We therefore determined the flux of Mab using a flux bootstrapping procedure similar to that applied to Keck-observed point sources by many other authors \citep[e.g.,][]{gibbard05, molter19, paradis23}, the specifics of which we summarize in Appendix \ref{s:bootstrapping}. We find the total flux of Mab to be $(5.2 \pm 1.7)\times10^{-16}$ \ergsec in H-band, where the error bars indicate a 1-$\sigma$ uncertainty, computed by taking the quadrature sum of the flux bootstrapping error (see Appendix \ref{s:bootstrapping}) and the 20\% photometric calibration error. This flux level corresponds to an integrated I/F of XX km$^2$ at 1.6 $\mu$m, where the integrated I/F is defined as the sky-projected area the moon would have if it were a perfect Lambertian reflector \citep[i.e., a geometric albedo of unity][]{karkoschka01}. Note that the Sun-target-observer phase angle of these observations was only 0.02$^\circ$, so the opposition effect was likely substantial, as discussed more in Section \ref{s:discussion}.

The observed position of Mab is the same as expected based on the orbital parameters in the JPL Horizons system\footnote{\url{https://ssd.jpl.nasa.gov/horizons/app.html\#/}} to within one Keck resolution element; see Figure \ref{fig:astrometry}. We determined this based on the relative positions of Mab, Miranda, and Puck, the latter two of which can be seen easily in a single 120-second frame. The advantage of using the positions of other moons as a reference for astrometry is that it does not depend on a navigation solution for the planet's disk.

\begin{figure}
\includegraphics[width=0.5\textwidth]{figures/expected_position_urh_Mab_oneframe.png}
\includegraphics[width=0.5\textwidth]{figures/expected_position_urh_Mab.png}
\caption{Two visualizations of Mab's expected position. \textbf{Left:} A single NIRC2 H-band exposure overlaid with the expected relative positions of Mab, Miranda, and Puck (red arrows) as well as the detected position of Mab (blue dot). The expected vector of motion of Mab is also overlain (blue arrow). \textbf{Right:} The same shifted-and-stacked detection image of Mab shown in the right panel of Figure \ref{fig:detection}, but zoomed out and stretched to show the apparent trajectories of Puck and Miranda. Mab's position is marked with a purple circle. The expected motion of Puck and Miranda relative to Mab are plotted as red and orange lines, respectively. Miranda is still overexposed in this image stretch, so the brightest pixels in Miranda's track are highlighted in cyan.\label{fig:astrometry}}
\end{figure}

A very faint region of positive flux was observed at the expected position of Mab after stacking the K-band data (see Figure \ref{fig:detection}). This feature is not clearly distinguishable by eye from positive noise spikes elsewhere in the frame. The false positive rate experiment discussed in Appendix \ref{s:falsepositives} shows that a noise spike of similar brightness would appear within $\sim$50 pixels of the expected position $\sim$8\% of the time. However, our uncertainty in Mab's position is roughly one H-band PSF FWHM, or $\sim$4 pixels, because we detected Mab at the expected position in H-band. We therefore label this a tentative detection. Photometry was performed in an identical way as for the H-band data, and we found a flux of $(3.6 \pm 0.9)\times10^{-16}$ \ergsec.

\subsection{Photometry of Other Moons}

We report the detection of Ophelia in H- and K-band and Perdita in H-band [CHECK K-BAND AGAIN].
Something about Cordelia! do we actually see it?
tie in story about median subtraction of frames.
Try Cupid again.

\section{Discussion}
\label{s:discussion}

The photometric model of \citet{karkoschka01} defines the I/F of a moon as the product of four normalized functions and a constant:
\begin{equation}
	\label{eq:if}
	I/F(l, \alpha, \lambda) = f_1(l) f_2(\alpha) f_3(\lambda) f_4(\lambda) (I/F)_0
\end{equation}
where $f_1(l)$ is the lightcurve as a function of orbital longitude $l$ with an orbital average of unity, $f_2(\alpha)$ is the phase function scaled to unity at phase angle $\alpha = 0$, $f_3(\lambda)$ is the solar spectrum (the ``continuum'') at the observed wavelength $\lambda$ scaled to unity at 550 nm, $f_4(\lambda)$ is the band/continuum ratio, used to account for wavelength-dependent absorption features in the spectrum of the moon, and $(I/F)_0$ is the reflectivity at $\alpha = 0$ and $\lambda = 550$ nm, averaged over orbital longitude.

Use the above to contextualize Showalter observations, Sam's upper limit, and the observed brightness in H-band vs K-band.

Sam's upper limit

What does it say about mu ring?

Something perhaps about orbital stability of Mab

\begin{acknowledgements}

The data presented herein were obtained at the W. M. Keck Observatory, which is operated as a scientific partnership among the California Institute of Technology, the University of California and the National Aeronautics and Space Administration. The Observatory was made possible by the generous financial support of the W. M. Keck Foundation.

The authors wish to recognize and acknowledge the very significant cultural role and reverence that the summit of Maunakea has always had within the indigenous Hawaiian community.  We are most fortunate to have the opportunity to conduct observations from this mountain.

\end{acknowledgements}

\appendix

\section{Shift-and-Stack Procedure}
\label{s:shiftandstack}

To improve the SNR on Mab, individual 120-second frames were shifted according to the expected position of the moon relative to Uranus and then co-added; details of this procedure are presented here. The accuracy of the shift-and-stack procedure depends sensitively on the rotation angle of the detector, so we corrected for the 0.26$^\circ$ angular offset between NIRC2's sky North position and true sky North before starting \citep{service16}. First, we aligned the position of Uranus in all the frames using the \texttt{chi2\_shift} function of the \texttt{Astropy}-affiliated \texttt{image\_registration} Python package\footnote{\url{https://github.com/keflavich/image_registration}}, which implements the single-step discrete Fourier transform (DFT) algorithm for efficient sub-pixel image registration \citep{guizarsicairos08}. In K-band, cloud features and moons are brighter than the planet's disk, so we first applied a Canny edge-detection filter using the \texttt{skimage} Python package \citep{skimage14} to enhance the rings of Uranus, then ran \texttt{chi2\_shift} on those edges. This alignment method was found to produce very sharp stacked images of the rings, so we were confident that it was aligning the frames accurately. The implementation of this alignment step permitted us to co-add all three H-band observing blocks (and two K-band observing blocks) shown in Figure \ref{fig:detection}, between which Uranus was repositioned on the detector by tens of pixels, and is the most important difference between our procedure and that of \citet{paradis19, paradis23}.  Second, we found the median of these planet-aligned frames and subtracted it from each frame; this step suppresses the signal of non-moving sources (in this case, the planet and rings), decreasing the effects of scattered light. Third, the expected offset between Mab and Uranus at the time of each exposure (we define this as the start time plus half the exposure time) was taken from JPL Horizons, accessed using the \texttt{Astroquery} Python package \citep{ginsburg19}. We converted the expected x,y offset from units of arcseconds to pixels assuming the pixel scale of the NIRC2 narrow camera to be 0.009971\arcsec \citep{service16} and taking the distance to Uranus from the Horizons ephemeris. Finally, we applied the shifts using a 2-D fast Fourier transform (FFT) sub-pixel image shift, again implemented with \texttt{image\_registration}, and summed all the frames.

\section{Flux Bootstrapping Technique}
\label{s:bootstrapping}

Using the \texttt{astropy}-affiliated \texttt{photutils} package \citep{photutils140}, we measured the total flux inside many circular apertures of different radii $r_i$ centered on Mab; let us call Mab's flux for the $i$th radius $F_{M i}$. We then measured the flux inside apertures of the same size but centered on the standard star; call these $F_{\star i}$. Comparing $F_{\star i}$ with the true flux of the standard star $F_\star$ (as measured in a very large aperture of radius $\gtrsim$100 pixels) yielded a correction for the wings of the PSF for a given value of $r_{i}$, given by simply $C_{psf} = F_{\star i}/F_\star$. The flux of the background must also be subtracted in order to accurately determine $F_{M i}$; we defined the background region using a circular annulus centered on Mab with inner radius $r_i + 5$ pixels and outer radius $r_o$. The background flux level $F_{B o}$ was taken to be the median of pixels within the background annulus multiplied by the area of the Mab aperture $\pi r_i^2$. We also varied $r_o$ over a range of values to test the sensitivity of our flux estimate to the choice of background region. The flux of Mab $F_{i,o}$ for a given $r_i, r_o$ pair is given by $F_M = (F_{M i} - F_{B o}) / C_{psf}$, and the value of $F_{i,o}$ is displayed as a function of $r_i$ and $r_o$ in Figure \ref{fig:wing}. The final flux estimate of Mab is given by the mean of $F_{i,o}$ over a reasonable range of $r_i$ and $r_o$ values, and the uncertainty is given by the standard deviation of $F_{i,o}$, which is $\approx$9\%. (This 9\% error is multiplied in quadrature with the typical 20\% flux calibration error from standard star observations with Keck). We chose smaller values of $r_i$ than \citet{paradis23} because the SNR of our Mab detection is much lower than the SNR of the moon detections shown in that paper.

\begin{figure}
\includegraphics[width=0.3\textwidth]{figures/example_aperture.png}
\includegraphics[width=0.7\textwidth]{figures/photometry_vs_region.png}
\caption{\textbf{Left:} Example aperture (solid blue line) and background annulus (dot-dashed red line) centered on Mab. The values of $r_i$ and $r_o$ in this example are 7 and 45 pixels, respectively. \textbf{Right:} Derived flux of Mab as a function of $r_i$ and $r_o$.\label{fig:wing}}
\end{figure}

\section{False Positive Rate Experiment}
\label{s:falsepositives}

Our detection of Mab has a low SNR, and the moon was located in the lower-left quadrant of the NIRC2 NARROW camera, which is known to have non-Gaussianity in its noise statistics,\footnote{https://www2.keck.hawaii.edu/inst/nirc2/ObserversManual.html} including vertical stripes of correlated noise. It is therefore natural to wonder whether the shift-and-stack technique is somehow injecting false positives by stacking noise spikes atop one another. To test whether this is the case, we performed an experiment to test the false positive rate within the true detector noise and scattered light environment of the data, as follows. The expected position of Mab was shifted by a random x,y offset of mean amplitude 20 pixels in each frame, then the shift-and-stack algorithm was re-run assuming these new randomly-selected shifts and the output frame was saved. This experiment was repeated 100 times with different random offsets applied. We then searched for point sources in these 100 test stacks by applying a center-surround (difference-of-Gaussains) filter with an FWHM of 4 pixels for the inner Gaussian\footnote{this corresponds roughly to the diffraction-limited beam size of Keck in H-band} and an FWHM of 12 pixels for the outer Gaussian. The center-surround filter is the appropriate choice because it accentuates features at the desired spatial frequency and downweights low spatial frequency features. The visual systems of humans and other animals employ center-surround antagonism \citep{graham06}, so this technique is a reasonable proxy for ``by-eye'' detection. We then binned the filtered images by a factor of 2 in each direction. Finally, we computed the amplitude of the brightest pixel in each binned frame in units of the RMS noise of the input image. This yielded a rough estimate of the signal-to-noise ratio of that peak. Identical filtering, binning, and SNR calculations were run on the real data. The detection of Mab in the real data was found to have a higher SNR than any peak in any of the 100 test frames in H-band (see Figure \ref{fig:randomstack}). In K-band, we find that a noise spike with an amplitude higher than our tentative detection is present in 8 out of 100 frames (Figure \ref{fig:randomstack_K}). However, because our tentative detection is in the correct position on the detector relative to Miranda and Puck, we feel justified in labeling it a tentative detection rather than a non-detection and extracting photometry. We tested other choices of inner and outer FWHM for the center-surround filter and different amounts of binning, and found that the results remained similar within reasonable ranges of these parameters. The SNR values reported in Figures \ref{fig:randomstack} and \ref{fig:randomstack_K} may look surprisingly high; note that they represent the detection level above the per-pixel RMS noise in the image, but these values do not correlate well with the true confidence level of a detection due to the substantial spatial correlation of the noise.

\begin{figure}
\includegraphics[width=1.0\textwidth]{figures/random_stack_experiment_Mab_H.png}
\caption{Results of the false positive rate experiment for H band. \textbf{(a)} Detection of Mab shown in Figure \ref{fig:detection}. \textbf{(b)} Panel (a) after filtering and binning were applied. \textbf{(c)} Example filtered and binned randomized stack. \textbf{(d)} Histogram of the SNR of the brightest pixel in the filtered and binned test frames; the real data are demarcated with a black dashed line.\label{fig:randomstack}}
\end{figure}

\begin{figure}
\includegraphics[width=1.0\textwidth]{figures/random_stack_experiment_Mab_K.png}
\caption{Results of the false positive rate experiment for K band. See caption of Figure \ref{fig:randomstack} for a description. \label{fig:randomstack_K}}
\end{figure}

\bibliography{references}

\end{document}